Converging High-Throughput and High-Performance Computing: A Case Study

### 2017-04-04 PandaEvolution.tex
Address Danila's and Alexei's comments

% as well as commercial and academic clouds.
% \mtnote{Addresses Alexei's comment. Increased Grid capacity to 300K cores. Is
% the number of sites still 100 or should I increase that too?}

% -----------------------------------------------------------------------------
% \subsection{Use of Supercomputers with PanDA}
% \label{ssec:panda-supercomputers}

@@

% \sergeynote{specifically "distributed Grid computing"}\mtnote{done.}.
% \sergeynote{geographically distributed?}\mtnote{Better?}

@@

% \mtnote{Move this to the introduction: ``There are at least two approaches to
% enable PanDA WMS to execute ATLAS workloads or workflows on supercomputers: (i)
% using the subset of resources and capabilities shared by both supercomputers and
% WLCG; (ii) reconciling the parallel and distributed computing paradigms by means
% of dedicated abstractions. The former is a pragmatic approach that enables the
% execution of specific workloads by prototyping single-point solutions. The
% latter is a principled approach, better suited for a production-grade solution,
% capable of supporting general-purpose workloads and workflows on both
% supercomputers and Grid infrastructures. \\ This section illustrates the design
% and architecture of a job broker prototyped by the PanDA team. This broker
% supports execution of part of the ATLAS production Monte Carlo workflow on
% Titan, a leadership-class supercomputer managed by the Oak Ridge Leadership
% Computing Facility (OLCF) at the Oak Ridge National Laboratory (ORNL).After an
% analysis of the results obtained and the lessons learned, the following section
% introduces the design and first experimental characterization of a next
% generation executor (NGE). NGE is designed to abstract resources and
% capabilities, enabling the concurrent execution of both parallel and distributed
% computing on generic high performance computing (HPC) machines.''}
%
% \jhanote{Do you mean the top of section 3? Or do you mean the very beginning
% section?}

% -----------------------------------------------------------------------------
%\subsection{Interfacing PanDA with Titan}
\subsection{Architectures and Interfaces}
\label{ssec:panda-titan}

@@

% Titan, was the first large-scale system to use a hybrid architecture that
% utilizes worker nodes with both AMD 16-core Opteron 6274 CPUs and NVIDIA Tesla
% K20 GPU accelerators. This hybrid design provides improved energy efficiency,
% as well as an order of magnitude in computational capacity over its
% predecessor.

@@

% Data staging is supported via the DTNs, from the wide area network and from
% the OLCF high performance storage system (HPSS). DTNs are shared across all
% Titan's users that can schedule jobs to a dedicates batch system to automate
% data staging. Fair-share policies are in place both for the batch system and
 %for the use of DTN resources.

@@

% \mtnote{Addresses Alexei's comment. Is the following sentence correct?}

@@

% While this is different from the HEP computing paradigm, where jobs are
% independent, it still shares common features such as the use of
% parallelization. It is not a requirement that HPC machines are able to run
% any possible task, nor is it relevant how many kinds of job types that can be
% run. What matters is the total number of cycles that can be offloaded from
% the Grid.

% Standard ATLAS workflow can not be easily ported to supercomputers due to
% several complications such as specialized worker node setups, no outbound
% network connections, limited memory per node, custom operating systems, etc. A
% reorganization of the standard workflow is therefore needed.

% -----------------------------------------------------------------------------
% \subsection{PanDA Broker on Titan}
\subsection{PanDA Broker}
\label{ssec:panda_titan}

@@

% MC event generation and simulation tasks can be executed without pilots but
% the time spent waiting in the queue would be too large compared to the
% execution time of each task.

% Titan’s backfill functionality offers the opportunity to avoid the overhead of
% queue wait times without using pilot abstraction. Backfill availability is the
% number of node-hours (or core-hours) that cannot be used for a certain amount
% of time by any of the jobs already queued on Titan: All queued jobs are either
% too large or their walltime is too long.

@@

% It is important to note that backfill availability is computed retrospectively
% and is determined by Titan's operation team. Backfill availability represents
% the theoretical maximum of node-hours that could be utilized by ATLAS (or any
% other project) that submit under backfill mode.

@@

% \sergeynote{any other job currently in the Titan's queue?}\mtnote{Better?}.

At any point in time, the size of an available backfill slot is typically a
small fraction of the total capacity of a resource. Notwithstanding, given the
size of Titan this translates into a substantial capacity. Every year, about
10\% of Titan's capacity remains unused~\cite{barker2016us}, corresponding to an
average of 30,000 unused cores (excluding GPU cores). This equals
% approximately 270M core-hours per year,
to roughly 10\% of the overall capacity of WLCG.

% A system that can use those temporarily free worker nodes with smaller
% scale tasks would be very valuable. For this reason, Titan offers backfill
% functionalities: users can interrogate Titan's Moab scheduler about how many
% free nodes are currently available and for how long. If a user submit a
% requestvfor that number of nodes and walltime, queue time will be minimal.

% \mtnote{Please feel free to add details about backfill as needed.}
% \mtnote{Please note: I know the following departs for the current name given
% to the PanDA subsystem installed on Titan. I think there is a good reason to
% call it a broker instead of a pilot, and I think I explained it in the
% previous paragraphs. Please  take this just as a suggestion, something I would
% like to discuss in our meeting.} \sergeynote{It's not a pilot in conventional
% sense. I call it an agent. It just happened that we were able to use full
% PanDA pilot's code base to serve our purposes on Titan. that's why, by
% inertia, we still call it a pilot.}\mtnote{Thank you. We would have a problem
% calling it `agent' as we use that term to name the pilot of NGE. Would PanDA
% Broker work?}\sergeynote{OK broker it is for this paper.}

Given the communication requirements of PanDA Pilots and the unused capacity of
Titan, PanDA pilot was repurposed to serve as a job broker on the DTN nodes of
Titan (Fig.~\ref{fig:panda_broker}). Maintaining the core modules of PanDA Pilot
and its stand-alone architecture, this prototype called `PanDA Broker'
implements functionalities to: (i) interrogate Titan about backfill
availability; (ii) pull MC jobs and events; (iii) wrap the payload of ATLAS jobs
into MPI scripts; (iv) submitting MPI scripts to Titan's PBS batch system and
monitor their execution; and (v) staging input/output files. Backfill querying,
payload wrapping, and scripts submission required a new implementation while
pulling ATLAS job and events, and file staging were inherited from PanDA Pilot.

% \sergeynote{Monitor job state and state transitions using
% SAGA}\mtnote{Better?}

@@

% Each script knows its MPI rank (an index that runs from zero to a maximum
% number of nodes or script copies) as well as the total number of ranks in the
% current submission. When activated on worker nodes, each copy of the MPI
% script

@@

The data staging capabilities of the PanDA Broker are implemented via a file
system that is shared among DTNs and worker nodes. The input files with the
events of the ATLAS jobs are downloaded on the shared filesystem via the ATLAS
DDM service. The MPI script setup process includes making the location of these
files available to the payload of the ATLAS's jobs. The PanDA Broker can locate
the payload's output files on the shared filesystem and transfer them from Titan
to any computing center used by ATLAS.
% the ATLAS Tier 1 computing center at Brookhaven National Lab (BNL) or any
% other Grid site.

Once deployed on Titan, every PanDA Broker supports the execution of MC detector
simulation in 9 steps. PanDA Broker queries the Job Dispatcher module of the
PanDA server for ATLAS jobs that have been bound to Titan by JEDI
(Fig.~\ref{fig:panda_broker}:1). Upon receiving jobs descriptions, PanDA Broker
pulls jobs' input files from the ATLAS DDM service to the OLCF Lustre file
system (Fig.~\ref{fig:panda_broker}:3). PanDA Broker queries Titan's Moab
scheduler about the current available backfill slot
(Fig.~\ref{fig:panda_broker}:2) and creates an MPI script, wrapping enough ATLAS
jobs' payload to fit the backfill slot. PanDA Broker submits the MPI script to
the Titan's Torque batch system via RADICAL-SAGA
(Fig.~\ref{fig:panda_broker}:4).

Upon execution on the worker node(s) (Fig.~\ref{fig:panda_broker}:5), the MPI
script initializes and configures the execution environment
(Fig.~\ref{fig:panda_broker}:6), and executes one AthenaMP for each available
work node (Fig.~\ref{fig:panda_broker}:7). AthenaMP retrieves events from Lustre
(Fig.~\ref{fig:panda_broker}:8) and spawns 1 event simulation process on each of
the 16 available cores (Fig.~\ref{fig:panda_broker}:9).
% During execution, PanDA Broker monitors execution progress and sends PanDA
% Server ``heart beats'' for each job.; and (viii)
Upon completion of each MPI script, PanDA Broker transfer the jobs' output to
designated computing centre via the ATLAS DDM(Fig.~\ref{fig:panda_broker}:10),
and performs cleanup.

% \mtnote{Address Alexei's comments: Add a paragraph explaining how this solution
% has been ported to other HPC machines and reference relevant publications.}
While PanDA Broker implementation is resource specific, it was successfully
ported also to other supercomputers, including the Ansel (?) at \ldots and
Cori at the National Energy Research Scientific Computing Center (NERSC).

% from the Job Dispatcher module of PanDA Server (as done by a PanDA Pilot);
% (iii) downloading events via PanDA Server's Data Service from ATLAS DDM to
% Titan's DTN; (iv) wrap the Geant4 payload into a PBS job; (v) submit jobs to
% Titan's PBS batch scheduler;

% Integration with Titan is the current focus for PanDA developers.

% The project aims to integrate Titan with the PanDA system using an updated
% PanDA Pilot that runs on the front-end node and submits ATLAS payloads to the
% worker nodes using the local batch system (PBS) via the SAGA (Simple API for
% Grid Applications) interface~\cite{SAGA}. This solves several complications
% of running on HPC worker nodes, including the lack of connectivity to outside
% world. The pilot can communicate with the PanDA server from the front-end
% machine.

% \paragraph*{HPC Backfill} HPC facilities are geared towards large scale jobs
% by design. Time allocation on an HPC is competitive and large projects are
% often preferred. About 10\% of capacity on a typical HPC machine is unused
% due to mismatches between job sizes and available resources. The worker nodes
% sit idle because there are not enough of them or they do not have enough time
% to handle a large scale computing job. On a machine of the scale of Titan,
% these 10\% correspond to estimated 300M core hours per year. A system that
% can occupy those temporarily free nodes with smaller scale tasks would be
% very valuable.

% This offers a great possibility for PanDA to harvest the opportunistic
% resources on Titan and use a similar mechanism on other HPCs. Functionality
% has been added to the PanDA Pilot to collect information about available
% unused worker nodes on Titan in real time.

% First test of the system were quite successful and show great promise in
% increasing the resource utilization on Titan.

>>>>

@@

@@

% -----------------------------------------------------------------------------
\subsection{Architectures and Interfaces}
\label{ssec:panda-titan}

@@

@@

@@

@@

% -----------------------------------------------------------------------------
\subsection{PanDA Broker}
\label{ssec:panda_titan}

@@

@@

@@

At any point in time, the size of an available backfill slot is typically a
small fraction of the total capacity of a resource. Notwithstanding, given the
size of Titan this translates into a substantial capacity. Every year, about
10\% of Titan's capacity remains unused~\cite{barker2016us}, corresponding to an
average of 30,000 unused cores (excluding GPU cores). This equals to roughly
10\% of the overall capacity of WLCG.

Given the communication requirements of PanDA Pilots and the unused capacity of
Titan, PanDA pilot was repurposed to serve as a job broker on the DTN nodes of
Titan (Fig.~\ref{fig:panda_broker}). Maintaining the core modules of PanDA Pilot
and its stand-alone architecture, this prototype called `PanDA Broker'
implements functionalities to: (i) interrogate Titan about backfill
availability; (ii) pull ATLAS jobs and events from PanDA Server; (iii) wrap the
payload of ATLAS jobs into MPI scripts; (iv) submitting MPI scripts to Titan's
PBS batch system and monitor their execution; and (v) staging and preparing
input/output files. Backfill querying, payload wrapping, and scripts submission
required a new implementation while pulling ATLAS job and events, and file
staging were inherited from PanDA Pilot.

@@

@@

The data staging capabilities of the PanDA Broker are implemented via a file
system that is shared among DTNs and worker nodes. The input files with the
events of the ATLAS jobs are downloaded on the shared filesystem from the data
center of Brookhaven National Laboratory (BNL). The MPI script setup process
includes making the location of these files available to the payload of the
ATLAS's jobs. The PanDA Broker can locate the payload's output files on the
shared filesystem and transfer them from Titan BNL.

Once deployed on Titan, every PanDA Broker supports the execution of MC detector
simulations in 9 steps. PanDA Broker queries the Job Dispatcher module of the
PanDA server for ATLAS jobs that have been bound to Titan by JEDI
(Fig.~\ref{fig:panda_broker}:1). Upon receiving jobs descriptions, PanDA Broker
pulls jobs' input files from BNL to the OLCF Lustre file system
(Fig.~\ref{fig:panda_broker}:2). PanDA Broker queries Titan's Moab scheduler
about the current available backfill slot (Fig.~\ref{fig:panda_broker}:3) and
creates an MPI script, wrapping enough ATLAS jobs' payload to fit the backfill
slot. PanDA Broker submits the MPI script to the Titan's Torque batch system via
RADICAL-SAGA (Fig.~\ref{fig:panda_broker}:4).

Upon execution on the worker node(s) (Fig.~\ref{fig:panda_broker}:5), the MPI
script initializes and configures the execution environment
(Fig.~\ref{fig:panda_broker}:6), and executes one AthenaMP for each available
work node (Fig.~\ref{fig:panda_broker}:7). AthenaMP retrieves events from Lustre
(Fig.~\ref{fig:panda_broker}:8) and spawns 1 Geant4 event simulation process on
each of the 16 available cores (Fig.~\ref{fig:panda_broker}:9). Upon completion
of each MPI script, PanDA Broker transfer the jobs' output to BNL
(Fig.~\ref{fig:panda_broker}:10), and performs cleanup.

While PanDA Broker implementation is resource specific, it was successfully
ported to other supercomputers, including the HPC2 at the National Research
Center ``Kurchatov Institute'' (NRC-KI)~\cite{belyaev2015integration},
Edison/Cori at the National Energy Research Scientific Computing Center
(NERSC)~\cite{barreiro2016panda}, and SuperMUC at the Leibniz Supercomputing
Centre (LRZ)~\cite{barreiro2016panda}.
